Self-tunable approximated explicit MPC: Heat exchanger implementation and analysys

### 2023-01-27 paper/elsarticle/main.tex
Review: Sec.2.2

[TODO: Please, write a paper structure]

@@

\textcolor{red}{TODO: Please, fix}

@@

The aggressivity of the controller and whole nature of the control can be influenced by tuning the penalty matrices in the optimization problem~\eqref{eq:mpc_problem}. When the optimization problem is precomputed offline to obtain the parametric solution, it is not possible to tune the controller afterwards. As the operation conditions and demands on controller setup may differ throughout control, the ability to adjust the controller aggressivity can be very beneficial.

The idea of approximated tunable explicit MPC comes from the work ~\cite{Klauco_tunable}, where the control action is calculated based on linear interpolation between two boundary control actions. These control actions result from evaluating two boundary explicit MPCs. The boundary explicit controllers have the same structure and setup, except one of the penalty matrices -- the tuned one. The boundary matrices are diagonal square matrices such that $\lambda_{i,\mathrm{L}} \le \lambda_{i,\mathrm{U}}$, $\forall i = 1,\dots,s$, where $\lambda$ denotes the vector of penalty matrix eigenvalues, $L$ and $U$ denote the lower and upper boundary setup respectively, and $s$ is the rank of the tuned penalty matrix.

When considering the penalty matrices from cost function~\eqref{eq:mpc_problem_cost}, the penalty matrices are scaled in the following way:
\begin{subequations}
	\label{eq:tunable_matrices}
	\begin{eqnarray}
		\label{eq:tunable_R}
		&~& R = (1-\rho)R_\mathrm{L} + \rho R_\mathrm{U}, \\
		\label{eq:tunable_Qx}
		&~& Q_\mathrm{x} = (1-\rho)Q_\mathrm{x,L} + \rho Q_\mathrm{x,U}, \\
		\label{eq:tunable_Qy}
		&~& Q_\mathrm{y} = (1-\rho)Q_\mathrm{y,L} + \rho Q_\mathrm{y,U},
	\end{eqnarray}
\end{subequations}

where $\rho$ represents the tuning parameter and $0 \le \rho \le 1$ holds. Based on~\eqref{eq:tunable_matrices}, it is possible to choose online any controller setup from the lower to the upper boundary of the tuned matrix. Note, it is practical to tune only one matrix, i.e., to store only two controllers corresponding to the boundary matrices. To determine which penalty matrix should be tuned, it is suggested to examine the control results subject to the specific system and to try systematic tuning of all the penalty matrices.

When the tuning parameter $\rho$ is determined based on the current control conditions, the optimal control action is evaluated using the two optimal controllers. Based on the boundary control actions, the interpolated, i.e., tuned control action is calculated:
\begin{subequations}
	\begin{eqnarray}
		\label{eq:tunable_u}
		u = (1-\rho)u_\mathrm{L} + \rho u_\mathrm{U},
	\end{eqnarray}
\end{subequations}

where $u_\mathrm{L}$ and $u_\mathrm{U}$ denote the optimal control actions from the lower and upper boundary controller respectively. The tuning of the controller online comes with a cost of storing and evaluating two explicit controllers. Nevertheless, the ability to tune the controller may be more important in many practical applications.

Note, the concept of explicit MPC tuning can be applied also to other MPC formulations, based on the current specific needs. In this paper, the penalty matrices of cost function~\eqref{eq:mpc_problem_cost} are considered, as it is necessary to satisfy offset-free reference tracking.

\begin{remark}
	If stability guarantee and recursive feasibility is required, it is suggested to follow the instructions from \cite{Oravec_tunable}. In order to satisfy the requirements, the study introduces a procedure of choosing the terminal penalty and terminal set in the two boundary controllers. 
\end{remark}

@@

\section*{Acknowledgements}
This research is funded by the European Union’s Horizon Europe under grant no. 101079342 (Fostering Opportunities Towards Slovak Excellence in Advanced Control for Smart Industries). The authors gratefully acknowledge the contribution of the Scientific Grant Agency of the Slovak Republic under the grants 1/0545/20, 1/0297/22, the Slovak Research and Development Agency under the project APVV-20-0261.

>>>>

\textcolor{red}{[TODO: Please, clearly point out the main contributions]}

This work directly extends our results presented in~\cite{self_tunable}, where the basic principles of the self-tunable approximated explicit MPC were introduced. Compared to~\cite{self_tunable}, we...

\textcolor{red}{[TODO: Please, write a paper structure]}

@@

\textcolor{red}{[TODO: Please, fix.]}

@@

The aggressivity of the controller and whole nature of the control is influenced by appropriate fine tuning of the penalty matrices in the optimization problem~\eqref{eq:mpc_problem}. When the multi-parametric QP (mp-QP) problem is precomputed offline to obtain the corresponding parametric solution, it is not possible to tune the controller afterwards without trading off a significant increase of the controller complexity or the performance loss. As the operation conditions and requirementson controller setup may differ throughout control, the ability to adjust the controller aggressivity can be very beneficial.

The idea of approximated tunable explicit MPC comes from the work ~\cite{Klauco_tunable}, where the control action is calculated based on linear interpolation between two boundary control actions. These control actions result from evaluating two boundary explicit MPCs. The boundary explicit controllers have the same structure and setup, except one of the penalty matrices -- the tuned one. The boundary penalty matrices are diagonal square matrices such that $\lambda_{i,\mathrm{L}} \le \lambda_{i,\mathrm{U}}$, $\forall i = 1,\dots,s$, where $\lambda$ denotes the vector of eigenvalues of the penalty matrix, $s$ is the rank of the tuned penalty matrix, and $L$, $U$ denote the lower and upper boundary setup respectively.

When considering the penalty matrices in cost function~\eqref{eq:mpc_problem_cost}, these penalty matrices are scaled in the following way:
\begin{subequations}
	\label{eq:tunable_matrices}
	\begin{eqnarray}
		\label{eq:tunable_R}
		&~& R(t) = (1-\rho(t)) \, R_\mathrm{L} + \rho(t) \, R_\mathrm{U}, \\
		\label{eq:tunable_Qx}
		&~& Q_\mathrm{x}(t) = (1-\rho(t)) \, Q_\mathrm{x,L} + \rho(t) \, Q_\mathrm{x,U}, \\
		\label{eq:tunable_Qy}
		&~& Q_\mathrm{y}(t) = (1-\rho(t)) \, Q_\mathrm{y,L} + \rho(t) \, Q_\mathrm{y,U},
	\end{eqnarray}
\end{subequations}
where $\rho$ represents the tuning parameter such that $0 \le \rho \le 1$ holds. Based on the rules in~\eqref{eq:tunable_matrices}, it is possible to choose online any controller setup from the lower to the upper boundary of the tuned matrix. From the implementation point of view, it is preferred to tune just a single penalty matrix, i.e., to store only two controllers corresponding to the boundary values of the selected penalty matrix. To determine which penalty matrix in~\eqref{eq:tunable_matrices} should be tuned, it is suggested to judge the control performance subject to the specific system and systematically tune all of the penalty matrices.

When the tuning parameter $\rho$ is determined based on the current control conditions, the approximated optimal control action is evaluated using the two optimal controllers. Based on the boundary control actions, the interpolated, i.e., tuned control action is calculated using the convex combination:
%\begin{subequations}
	\begin{eqnarray}
		\label{eq:tunable_u}
		u(t) = (1-\rho(t)) \, u_\mathrm{L} + \rho(t) \, u_\mathrm{U},
	\end{eqnarray}
%\end{subequations}
where $u_\mathrm{L}$ and $u_\mathrm{U}$ denote the optimal control actions from the lower and upper boundary controller respectively. The online tuning of the controller online comes with a cost of storing and evaluating two explicit controllers. Nevertheless, the ability to tune the controller may be more important in many practical applications.

We point out, the concept of explicit MPC tuning is applicable for a wide class of MPC design formulations, based on the current specific needs. Without loss of generality, hereafter, we consider the penalty matrices of cost function~\eqref{eq:mpc_problem_cost}, as it is necessary to satisfy offset-free reference tracking.

\begin{remark}
	If the asymptotic stability and recursive feasibility guarantees are required, we refer the reader to follow the instructions from \cite{Oravec_tunable}. In order to satisfy these requirements, the study introduces a procedure of computing the common terminal penalty and terminal set for the two boundary controllers. 
\end{remark}

@@

\section*{Acknowledgements}

The authors gratefully acknowledge the contribution of the Scientific Grant Agency of the Slovak Republic under the grants 1/0545/20, 1/0297/22, the Slovak Research and Development Agency under the project APVV-20-0261. 
This research is funded by the European Union’s Horizon Europe under grant no. 101079342 (Fostering Opportunities Towards Slovak Excellence in Advanced Control for Smart Industries).
